### 2015-05-06 version1.0/application.tex
adding ms

%% APPLICATION
%%

We have presented a HBM to infer the probability distribution of a star's
surface gravity and bulk density using flicker. These models are calibrated by a 
comparison of flicker to asteroseismic values. In this section, we provide a
comparison of the predicted $\rho_{\star}$ values from our model versus an
independent metric, in order to validate our predictictions.

A suitable comparison comes from stars with transiting planets, where the shape 
of the transit light curve may also be used to derive $\rho_{\star}$. Several
effects can change