### 2015-05-12 version1.0/introduction.tex
adding words

%% INTRODUCTION
%%

% Stellar characterization techniaues overview, traditionally frequentist

% who cares about characterization?

Accurate stellar characterization plays a vital role for many active research 
fields within astronomy. For example, stellar populations, galactic archaeology, 
the study of binary stars, asteroseismology and exoplanet studies all rely on 
inferences of basic stellar parameters to varying degrees. Empirically-derived
and reliable estimates are of particular value, increasing our confidence in the
end-product results built upon these inputs.

Basic stellar parameters, such as effective temperature and surface gravity, 
can be inferred using one (or more) of several types of observations, e.g. 
spectroscopy, photometry, interferometry, etc. This inference can be done by
invoking theoretical models or by building an empirical calibration library.
As an example, an observed stellar spectrum could be matched against a library of 
theoretical spectra generated using stellar atmosphere models, or, against a 
library of observed spectra of ``standard stars'', serving as calibrators.
%Fundamental stellar parameters, such as mass and radius, which cannot be 
%directly inferred from the available observations can often be inferred by 
%invoking theoretical stellar models, e.g. fitting model isochrones to 
%spectroscopic measurements.
Regardless of the approach, be it theoretical or empirical, the methods used
for the inference of stellar parameters are traditionally ``deterministic''.
In this context, a deterministic model can be loosely described as one where
a particular observational input always returns a single-valued output for a 
parameter of interest. In this framework, nature itself has no variance and
our underlying model is considered to be a perfect description of reality.

An alternative approach for inferring model parameters is the 
``probabilistic'' one. Unlike the deterministic case, a single observational 
input is intepretted to be caused by a range of possible model parameters
described by a probability distribution. This statement is true even in the case
of a perfect observation of infinite signal-to-noise, since our underlying 
model itself comes with uncertainty. In recent years, there has been a shift 
towards such methods in several areas of astronomy, particularly within the 
exoplanet community. For example, \citet{wolfgang:2015} considered that
the mass-radius relationship of exoplanets is probabilistic, since a particular
sized planet could be due to a range of planet masses, thanks to variances in
compositions, environment and other circumstances. These recent demonstrations
in exoplanetary science have prompted us to consider the need for treating the 
parent stars in the same probabilistic framework, with potential applications 
spanning many fields of astronomy.

The demand for probilistic stellar parameters is not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, but also... (ruth will continue this sentence)

%A number of recent exoplanet studies have used Hierarchical Bayesian Modeling
%(HBM) to model the dispersion, or intrinsic scatter in relations between
%astrophysical properties.
%For example, \citep{wolfgang:2015} use HBM to recalibrate the mass-radius
%relation for sub-Neptunes and infer the level of astrophysical dispersion