### 2015-06-08 version1.0/introduction.tex

%% INTRODUCTION
%%

Accurate stellar characterization plays a vital role for many active research 
fields within astronomy. For example, stellar populations, galactic archaeology, 
the study of binary stars, asteroseismology and exoplanet studies all rely on 
inferences of basic stellar parameters to varying degrees. Empirically-derived
and reliable estimates are of particular value, increasing our confidence in the
end-product results built upon these inputs.

Basic stellar parameters, such as effective temperature and surface gravity, 
can be inferred using one (or more) of several types of observations, such as
spectroscopy, photometry, interferometry, etc. This inference can be done by
invoking theoretical models or by building an empirical calibration library.
For example, an observed stellar spectrum could be matched against a library of 
theoretical spectra generated using stellar atmosphere models, or, against a 
library of observed spectra of ``standard stars'', serving as calibrators.
%Fundamental stellar parameters, such as mass and radius, which cannot be 
%directly inferred from the available observations can often be inferred by 
%invoking theoretical stellar models, e.g. fitting model isochrones to 
%spectroscopic measurements.
Regardless of the approach, be it theoretical or empirical, the methods used
for the inference of stellar parameters are traditionally ``deterministic''.
In this context, a deterministic model can be loosely described as one where
a particular observational input always returns a single-valued output for a 
parameter of interest. In this framework, nature itself has no variance and
our underlying model is considered to be a perfect description of reality.

An alternative approach for inferring model parameters is the 
``probabilistic'' one. Unlike the deterministic case, a single observational 
input is intepretted to be caused by a range of possible model parameters
described by a probability distribution. This statement is true even in the case
of a perfect observation of infinite signal-to-noise, since our underlying 
model itself comes with uncertainty. In recent years, there has been a shift 
towards such methods in several areas of astronomy, particularly within the 
exoplanet community. For example, \citet{wolfgang:2015} considered that
the mass-radius relationship of exoplanets is probabilistic, since a particular
sized planet could be due to a range of planet masses, cdue to unmodeled 
variances in compositions, environment and other complications. These recent 
demonstrations in exoplanetary science have prompted us to consider the need for 
treating the parent stars in the same probabilistic framework, with potential 
applications spanning many fields of astronomy.

\textbf{
The demand for probilistic stellar parameters is not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, but also... (ruth will continue this sentence)
}

%A number of recent exoplanet studies have used Hierarchical Bayesian Modeling
%(HBM) to model the dispersion, or intrinsic scatter in relations between
%astrophysical properties.
%For example, \citep{wolfgang:2015} use HBM to recalibrate the mass-radius
%relation for sub-Neptunes and infer the level of astrophysical dispersion

% RA text...

One of the more recent tools developed to characterize stars is known as
``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an
8-hour timescale (denoted as $F_8$) in a broad visible bandpass time series
photometric light curve, such as that from \textit{Kepler} or the upcoming TESS
mission. A more detailed account of the proceedure to calculate flicker is
described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker
displays a remarkable correlation to the asteroseismically determined parent
star surface gravities (\logg). Turning this around, the observation implies
that flicker can be used to blindly infer surface gravities at the level of
$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light
curves available through the array of exoplanet transit missions flying and
scheduled to launch.

\citet{cranmer:2014} demonstrated that models of stellar surface granulation are 
indeed reproduce a flicker effect in close agreement with that observed by
\citet{bastien:2013}, providing a physically-plausible explanation. Since
surface gravity is highly correlated to mean stellar density (\rhostar) on
evolutionary tracks, \citet{kipping:2014} showed that flicker can be also
be used to infer \rhostar, which is more useful for exoplanet transit
analysis \citep{seager:2003}.

Whether one calibrates flicker to \logg\ or \rhostar, there are several aspects
of the problem which are attractive for our purposes of a simple demonstration
of probabilistic inference of stellar parameters. Firstly, in log-log space the
relationship is very simple, appearing to be linear \citep{kipping:2014}. 
Secondly, there is a sufficiently large number of points in the sample (439 
stars) to constrain a population-based model. Thirdly, there is significant 
excess scatter around the best-fitting relation implying that a deterministic 
model is inadequate. This is not surprising given that granulation is a complex 
and messy process for which one should not expect any parametric model to 
provide a perfect description. For these reasons, we identify the calibration of 
flicker to \logg\ and \rhostar\ as a well-posed problem to first demonstrate 
probabilistic inference in the arena of stellar characterization.